### 2016-05-11 Docs/Communications/CCS2016/ccs2016.tex

\input{WPHeader.tex}

\title{Unveiling Patterns of Technological Innovation through Data-mining\\\bigskip
\textit{Communication Proposal, CCS 2016}
}
\author{\small\noun{Juste Raimbault}$^{1,2}$\\
\small(1) UMR CNRS 8504 Géographie-cités and (2) UMR-T IFSTTAR 9403 LVMT
}
%\date{15 octobre 2015}
\date{}

\maketitle

\justify

\pagenumbering{gobble}

%\vspace{-0.5cm}
\textbf{Keywords : }\textit{}

The understanding of patterns of technological innovation is crucial for both economical growth theories and practical application in research and developments. Yet a precise characterization of breakthrough inventions has not been fully investigated. We propose a large-scale data-mining approach on technological patents.

\bigskip

\bigskip

%%%%%%%%%%%%%%%%%%%%
%% Biblio
%%%%%%%%%%%%%%%%%%%%

%\begin{multicols}{2}

%\setstretch{0.3}
%\setlength{\parskip}{-0.4em}

%\bibliographystyle{apalike}
%\bibliography{/Users/Juste/Documents/ComplexSystems/CityNetwork/Biblio/Bibtex/CityNetwork}

%\end{multicols}

\end{document}