### 2018-11-27 Docs/Papers/Technical/overleafv2/S5_Text.tex

\input{si.tex}

\section*{S5 Text : Statistical definitions and derivations}
\subsubsection*{Likelihood expression}
We define $\mathcal{F}_i$ the filtration which corresponds to the time $t_i$. With this notation $\mathcal{L}( X | \mathcal{F}_{i-1})$ simply means the likelihood of $X$ conditioned on the past. We consider $\widehat{\theta}^{(z)}$ the MLE\footnote{Apparently, this MLE is a partial MLE, but we will not refer to partial for simplicity.} of $\theta^{(z)}$, where the corresponding log-likelihood of the model can be expressed up to constant terms as
\begin{eqnarray*}
\sum_{i=2}^Z \sum_{l=1}^{C_i} \log \mathcal{L} \big( B_{l,i} | \mathcal{F}_{i-1} \big).
\end{eqnarray*}
Recalling that $B_{l,i}$ are independent of each other and conditioned on the past follow Bernoulli variables 
$$B \Big( \min \Big\{ 1, \frac{N_{i}^{(z)}}{i-1} + \theta^{(z)} \Big\} \Big),$$ 
the log-likelihood of the model can be expressed as 
\begin{eqnarray}
\label{loglik}
\sum_{i=2}^Z \sum_{l=1}^{C_i} B_{l,i} \log \Big( \min \Big\{ 1, \frac{N_{i}^{(z)}}{i-1} + \theta^{(z)} \Big\} \Big) +(1 - B_{l,i}) \log \Big(1 - \min \Big\{ 1, \frac{N_{i}^{(z)}}{i-1} + \theta^{(z)} \Big\} \Big).
\end{eqnarray}
In practice, the user can easily implement the formula (\ref{loglik}) for any $0 \leq \theta^{(z)} \leq 1$, and maximize it over a predefined grid to obtain $\widehat{\theta}^{(z)}$.

\subsubsection*{Standard errors of the estimated values}
Under some assumptions, it is possible to show the asymptotic normality of $\widehat{\theta}^{(z)}$ and to compute the asymptotic variance. For simplicity of exposition, we assume that we restrict to $\theta^{(z)}$ such that we have $\frac{N_{i}^{(z)}}{i-1} + \theta^{(z)} < 1$ for any $i=2, \cdots, Z$. The central limit theorem can be expressed as 
\begin{eqnarray}
\label{variance}
\sqrt{Z \mathbb{E} [ C ]} (\widehat{\theta}^{(z)} - \theta^{(z)}) \overset{\mathcal{L}}{\rightarrow} MN \Big( 0, \int (p + \theta^{(z)})(1 - (p + \theta^{(z)})) d\pi^{(z)} (p) \Big),
\end{eqnarray}
where MN stands for a multinormal distribution and $\pi^{(z)}$ for the asymptotic limit distribution of the quantity $\frac{N_{i}^{(z)}}{i-1} + \theta^{(z)}$. Note that the variance term in (\ref{variance}) is equal to an aggregate version of the Fisher information matrix. The proof of such statement is beyond the scope of this paper. On the basis of (\ref{variance}), we provide a variance estimator as 
$$v^{(z)} = \frac{1}{C_k-1}\sum_{i=2}^{C_k}  \frac{N_{i_k}^{(z)}}{i-1} + \widehat{\theta}^{(z)},$$ 
where $i_k$ is such that the $i_k$th patent corresponds to the $k$th couple. This estimator was used to compute the standard deviation in Table \ref{summary}.
\subsubsection*{Test statistic}
The test statistic used is a mean difference test statistic between $\widehat{\theta}^{(tec)}$ and  $\widehat{\theta}^{(sem)}$, where the formal expression can be found in (\ref{teststat}). We assume independence between both quantities and thus under the null hypothesis, we have that 
$$\widehat{\theta}^{(tec)} - \widehat{\theta}^{(sem)} {\rightarrow} MN(0,V),$$
where $V= \int (p + \theta^{(tec)})(1 - (p + \theta^{(tec)})) d\pi^{(tec)} (p) + \int (p + \theta^{(sem)})(1 - (p + \theta^{(sem)})) d\pi^{(sem)} (p)$ can be estimated by $\widehat{V} = v^{(sem)} + v^{(tec)}$. Then, we obtain that 
\begin{eqnarray}
\label{teststat}
A = \frac{\widehat{\theta}^{(tec)} - \widehat{\theta}^{(sem)}}{\widehat{V}} \approx \mathcal{N} (0,1),
\end{eqnarray}
where $A$ is the mean difference test static.

\end{document}